---
title: ".Random.seed"
output:
  pdf_document: default
  html_document: default
---

## Setting the seed

You will want to set the seed, or find out what value the seed has and save that value, before you begin generating your random nodes. `.Random.seed` does not exist until a random number has been generated or the seed has been set.  We set the seed and then save it with

```{r}
seeds = vector(length = 2, mode = "list")
set.seed(12345678)
seeds[[1]] = .Random.seed
```

Note that the saved seed is not the same as the value to `set.seed`.  `.Random.seed` is a vector of over 600 integers which are used by the random number generator, and the `set.seed` function creates these values based on the input given.


## Saving the seed as you progress

You may want to save the seed, before you create a new set of nodes.

For example, suppose that you run `runif` many times

```{r, collapse=TRUE}
x = runif(10)
x
invisible(replicate(17, runif(10)))
```

Next let's save the current value of the seed and generate more random values.

```{r, collapse=TRUE}
seeds[[2]] = .Random.seed
y = runif(7)
y
invisible(replicate(12, runif(3)))
```

Notice that we used `invisible` to keep all of the random values from printing in our document.


Now if you want to recreate `x`, we can set the seed back to `seeds[[1]]` with

```{r, collapse=TRUE}
.Random.seed = seeds[[1]]
x2 = runif(10)
x2
all(x2 == x)
```

Or if you want to recreate `y`, we can set the seed back to `seeds[[2]]` with

```{r,collapse=TRUE}
.Random.seed = seeds[[2]]
y2 = runif(7)
y2
all.equal(y2, y)
```

Here, we use the `all.equal` function to compare the values in `y` and `y2`.